---
title: "cohort_descriptive"
author: "Stefanie Mueller"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    lightbox: true
    gallery: true
    toc_depth: 3
    use_bookdown: true
    fig_caption: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r knitr_init, echo=FALSE, cache=FALSE}
# load libraries
library(knitr)
library(rmdformats)

# define functions
  myDT = function(data){
  data %>% DT::datatable(rownames = FALSE,
    filter = 'bottom',
    extensions = 'Buttons',
      options = list(
        searching = TRUE,
        fixedColumns = TRUE,
        autoWidth = TRUE,
        dom = 'Bfrtip',
        buttons = c('copy', 'csv', 'excel'))
    )
}

## Global options
#options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=TRUE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               fig.width = 10)
opts_knit$set(width=75)
```

# PREFACE

## PROJECT
burden analysis extendend gene region (including regulatory) in sporadic breast cancer in diverse ancestries

## OBJECTIVE
create cohort descriptive and some QC

## Of note
detailed phenotypes not shared in public github repository

## load libraries
```{r}
library(tidyverse)
library(data.table)
```

## Version Info
```{r}
sessionInfo()
```

## Set Global Varibales
```{r}
DIR="/home/stef/WORK/UCL/UGI/BCAC_publication"
OUTFOLDER=paste0(DIR, "data/output/")
INPUTFOLDER=paste0(DIR, "data/input/")
SCRIPT="S02"
DATE=Sys.Date()
```

## Plot Theme
```{r}
my_ggplotTheme = theme_bw()+
  theme(title = element_text(size=rel(1.3), face = "bold"),
        axis.title.y = element_text(vjust=1.5),
        axis.title.x = element_text(vjust=-1.5),
        axis.text = element_text(size=rel(1.3)),
        axis.text.x = element_text(hjust = 1),
        legend.text = element_text(size=rel(1.3)),
        strip.text = element_text(size=rel(1.3)),
        plot.margin = unit(c(1,1,1,2), "cm"),
        panel.grid.major = element_line(colour="grey60"))
```

# Load inputs

## Load phenos
```{r}
pheno_files = list.files(path = paste0(INPUTFOLDER,"pheno"),
                  pattern = "detailed_phenos.txt$",
                  full.names = T)

cohort = basename(pheno_files) %>% 
  str_split(pattern = "\\.") %>% 
  lapply("[",1) %>%  unlist()

pheno.list = list()

for (i in 1:length(pheno_files)){  
 
  pheno.list[[i]] = fread(pheno_files[i]) %>% 
    select(-starts_with("PC_"), -starts_with("study.")) %>% 
    mutate(COHORT=cohort[i]) 
}

pheno.DF = data.table::rbindlist(pheno.list)
```

## save long pheno data
```{r}
pheno.DF %>%   write_csv(paste0(OUTFOLDER, SCRIPT,".supp.all_phenos.",DATE,".csv"))
```

# Make Table1 
## Overall
```{r}
pheno.DF %>% 
  group_by(Pheno) %>%
  summarise(Count = n(),
            Age.median= median(Age, na.rm=T),
            Age.iqr = quantile(Age, probs = 0.75, na.rm=T) -quantile(Age, probs = 0.25, na.rm=T),
            African.prop = mean(EthnicityGeno=="African"),
            African.num = sum(EthnicityGeno=="African"),
            Asian.prop = mean(EthnicityGeno=="Asian"),
            Asian.num = sum(EthnicityGeno=="Asian"),
            European.prop = mean(EthnicityGeno=="European"),
            European.num = sum(EthnicityGeno=="European"),
            Hispanic.prop = mean(EthnicityGeno=="other"),
            Hispanic.num = sum(EthnicityGeno=="other")) %>% 
  write_csv(paste0(OUTFOLDER, SCRIPT,".supp.table1.",DATE,".csv"))
```

